---
title: 'Blog Post 01'
author: "Ben Sunshine"
date: "2/9/2024"
editor: visual
categories:
  - Sports
---

## Introduction

In this blog post, I am analyzing a data set of 2024 MLB Fantasy Baseball Projections. These projections are 'Zeile' Projections (sourced from [FantasyPros](https://www.fantasypros.com/mlb/projections/hitters.php)), which are baseball specific projections derived from a consensus of 7 sources including ESPN, Draft Buddy, Baseball Think Factory, Steamer Blog, Razzball, Derek Carty, and FanGraphs. This data set consists 747 observations (players) and 17 variables which include: 'Player', 'Team', 'Positions', 'AB', 'R', 'HR', 'RBI', 'SB', 'AVG', 'OBP', 'H', '2B', '3B', 'BB', 'SO', 'SLG', 'OPS'. This data set is hitter specific, so some of the variables I am most interested in include 'AVG' (Batting Average) and 'OPS' (On Base Percentage Plus Slugging Percentage). Using these variables, I aim to visualize which teams will have the strongest projected offenses.

```{r}
library(tidyverse)
projections <- read_csv('/Users/bensunshine/Documents/SLU_Senior_Year/SP24/data_334/ds334blog/data/FantasyPros_2024_Projections_H.csv')

missing_team <-
  projections %>%
  filter(is.na(Team))

num_players <- nrow(projections)

division_order <- c('AL EAST', 'AL CENTRAL', 'AL WEST', 'NL EAST', 'NL CENTRAL', 'NL WEST', NA)

projections <-
  projections %>%
  filter(!is.na(Team)) %>%
  mutate(Division = case_when(
    grepl('(BOS|NYY|TOR|BAL|TB)', Team) ~ 'AL EAST',
    grepl('(MIN|DET|CLE|CWS|KC)', Team) ~ 'AL CENTRAL',
    grepl('(HOU|TEX|SEA|LAA|OAK)', Team) ~ 'AL WEST',
    grepl('(ATL|PHI|MIA|NYM|WSH)', Team) ~ 'NL EAST',
    grepl('(MIL|CHC|CIN|PIT|STL)', Team) ~ 'NL CENTRAL',
    grepl('(LAD|ARI|SD|SF|COL)', Team) ~ 'NL WEST'
  )) %>%
  mutate(Division = factor(Division, levels = division_order))


```

```{r}
team_avg_projections <- 
  projections %>%
  group_by(Division, Team) %>%
  summarise(num_player = n(),
            avg_hr = mean(HR),
            avg_baa = mean(AVG),
            avg_hits = mean(H),
            avg_ops = mean(OPS),
            se_baa = sqrt(
              ((mean(AVG)/n()) * (1-(mean(AVG)/n()))) /
                n()
              )
            ) %>%
  mutate(Division = factor(Division, levels = division_order),
         lb_se_baa = avg_baa - se_baa,
         ub_se_baa = avg_baa + se_baa)
  
```

## Visualizations

```{r}
# error bar plot for mean BAA
team_avg_projections %>%
  mutate(Team = fct_reorder(Team, avg_baa), .desc = T) %>%
  ggplot(aes(x = Team, y = avg_baa)) +
  geom_errorbar(aes(ymin = lb_se_baa, ymax = ub_se_baa, colour = Division)) +
  geom_point(aes(x = Team, y = avg_baa)) +
  labs(title = "Mean BAA by Team",
       y = "Mean BAA") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = 'none') +
  facet_wrap(~ Division, scales = 'free_y') +
  coord_flip() +
  theme_minimal()
```

-   This first plot analyzes each team's mean batting average. Each team is grouped into a faceted section of the plot according to their division. The standard error is plotted in addition to the mean batting average for each team to display the variability and confidence intervals around the averages. It can be seen the Colorado Rockies are predicted to have highest average batting average in the 2024 season. Most teams have consistent standard errors, but it appears the Milwaukee Brewers have the smallest standard error, indicating their players are likely to have batting averages closer to their mean batting average than other teams.

```{r}
# bar plot mean OPS
team_avg_projections %>%
  mutate(Team = fct_reorder(Team, avg_ops), .desc = T) %>%
  ggplot(aes(x = Team,
             y = avg_ops,
             fill = Division)) +
  geom_col(color = 'black') +
  labs(y = 'Mean OPS',
       title = 'Mean OPS by Team') +
  scale_y_continuous(n.breaks = 6) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = 'none') +
  facet_wrap(~ Division, scales = 'free_y') +
  #coord_flip(ylim = c(0.22,0.26)) +
  coord_flip(ylim = c(0.6,0.75)) +
  theme_minimal()

```

-   The second plot examines each team's mean on base percentage plus slugging percentage (OPS) using bar plots faceted by their Division. This statistic is unique because it gives insight into the likelihood of a team getting on base and generating extra base hits. Again it can be the Colorado Rockies are projected to have the highest OPS in the MLB at over 0.720. Just behind are both the Boston Red Sox and the Atlanta Braves.

## Conclusion

Because there are 30 teams in the MLB, I decided to facet the teams by their division to make the plots seem less cluttered. In the first plot I also utilized the error bar for displaying the standard error as presented in class. This gives us more information about the distribution of batting averages for each team, which would be lost in the summarized data. To make both graphs easily interpretable, I also arranged each faceted plot to be sorted in descending order of the statistic being examined. This makes it easy for the viewer to see which team has the highest and lowest statistics in each division. In this data set it's important to note there were 89 players who had 'NA' values for their 'Team' variable. This is due to the fact that these players are free agents in 2024, so they are currently unassigned to a team. For simplicity, I removed these players from my analysis so I could compare teams and their current players. In the future, I would be interested in comparing each player's and team's projected statistics against their actual 2024 statistics. A huge amount is bet on fantasy sports, so it would be interesting to see how well these projections perform.
